```{r}
install.packages("dplyr")
install.packages("seastests")
install.packages("seasonal")
install.packages(c("tseries", "stats"))
```

```{r}
library(tseries)
library(stats)
library(seasonal)
library(seastests)
library(dplyr)
library(ggplot2)
library(readxl)
library(lubridate)
```

```{r}
data <- read_excel("C:\\Users\\vikas\\Desktop\\Term 3\\TS\\Project\\New_Cars.xlsx")

data$Month <- as.Date(data$Month, format = "%m/%d/%Y")

car_data <- filter(data, Month >= as.Date("2013-01-01"))
```

```{r}

# Plot the time series data
ggplot(car_data, aes(x = Month, y = Sales)) +
  geom_line() +
  labs(title = "Car Retail Sales Over Time",
       x = "Month",
       y = "Sales") +
  theme_minimal()

# Create a time series object
ts_data <- ts(car_data$Sales, frequency = 12)

# Decompose the time series using stl (Seasonal-Trend decomposition using Loess)
decomposition <- stl(ts_data, s.window = "periodic")

# Plot the decomposed components
plot(decomposition)

# Plot the seasonal component
plot(decomposition$time.series[, "seasonal"], main = "Seasonal Component")

# Plot the trend component
plot(decomposition$time.series[, "trend"], main = "Trend Component")

# Plot the remainder (residuals) component
plot(decomposition$time.series[, "remainder"], main = "Residuals (Remainder) Component")
```

An upward trend indicates a general increase in the variable over time. The lower bump in the middle of the trend suggests a temporary deviation or anomaly. This could be due to specific events, factors, or external influences affecting the underlying upward movement.

The deep groove in the residuals (remainder) where the trend graph has a bump indicates that the classical decomposition method might not fully capture the variation during that period. The residuals represent what is left after removing the trend and seasonal components. A deep groove suggests that there might be unexplained variation during that specific time, possibly due to outliers or irregularities.

The magnitude of the seasonal pattern does not change significantly throughout the entire time series.

```{r}
# Summary statistics
summary(ts_data)

# Augmented Dickey-Fuller (ADF) test
adf_test <- adf.test(ts_data)
print(adf_test)
```

The negative Dickey-Fuller test statistic and the low p-value provide evidence that your time series is likely stationary. The Augmented Dickey-Fuller test supports the hypothesis that there is no unit root in your data, suggesting stationarity.

```{r}
# Create a time series object (replace 'your_data' with your actual time series data)
ts_data <- ts(car_data$Sales, frequency = 12)

# Perform decomposition using X-11 method
decomposition <- decompose(ts_data, type = "additive")

# Print the components
print(decomposition)


```

```{r}
your_data <- ts_data
# Plot the original time series
plot(your_data, main = "Original Time Series")

# Plot the trend component
plot(decomposition)
```


```{r}

```
